---
title: "R Markdown for Course_Project_4"
author: "Charmaine Largo"
date: "May 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##R Markdown 
### Course Project 4

Here are the questions that need to be adressed for this project.

1. Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base
plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years
1999, 2002, 2005, and 2008.

2. Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips == "24510")
from 1999 to 2008? Use the base plotting system to make a plot answering this question.

3. Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable,
which of these four sources have seen decreases in emissions from 1999-2008 for Baltimore City?
Which have seen increases in emissions from 1999-2008?

4. Across the United States, how have emissions from coal combustion-related sources changed from
1999-2008?

5. How have emissions from motor vehicle sources changed from 1999-2008 in Baltimore City?

6. Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle
sources in Los Angeles County, California (fips == "06037"). Which city has seen greater
changes over time in motor vehicle emissions?


First, we need to load these the data sets.
```{r NEI}
NEI <- readRDS("summarySCC_PM25.rds")
```
```{r SCC}
SCC <- readRDS("Source_Classification_Code.rds")
```
## Question 1

First aggregate the Total PM2.5 emission for all the sources.
```{r total}
total <- aggregate(Emissions ~ year,NEI, sum)
```


```{r barplot}
barplot(
  (total$Emissions)/10^6,
  names.arg=total$year,
  xlab="Year",
  ylab="PM2.5 Emissions",
  main="Total PM2.5 Emissions From All US Sources"
)
```

The plot shows that there was a decrease in the US total emissions from 1999 to 2008.


## Question 2
Here, we need to aggregate the Total PM2.5 emission only for Baltimore City.
```{r baltimore}
baltimore <- NEI[NEI$fips=="24510",]
```

```{r totalBaltimore}
totalBaltimore <- aggregate(Emissions ~ year, baltimore,sum)
```

```{r barplot2}
barplot(
  totalBaltimore$Emissions,
  names.arg=totalBaltimore$year,
  xlab="Year",
  ylab="PM2.5 Emissions (Tons)",
  main="Total PM2.5 Emissions From All Baltimore City Sources"
)
```

There was a decrease in overall total emissions for Baltimore City from 1999 to 2008.

## Question 3

Call ggplot2 package.
```{r ggplot2}
library(ggplot2)
ggp <- ggplot(baltimore,aes(factor(year),Emissions,fill=type)) +
  geom_bar(stat="identity") +
  theme_bw() + guides(fill=FALSE)+
  facet_grid(.~type,scales = "free",space="free") + 
  labs(x="year", y=expression("Total PM"[2.5]*" Emission (Tons)")) + 
  labs(title=expression("PM"[2.5]*" Emissions, Baltimore City 1999-2008 by Source Type"))
```

Print the plot.
```{r print}
print(ggp)
```
Based on the plot non-road, nonpoint, on-road source types have all seen decreased emissions overall in Baltimore City from 1999-2008.

## Question 4

Subset the source factors of coal combustion using NEI data.
```{r combustionRelated}
combustionRelated <- grepl("comb", SCC$SCC.Level.One, ignore.case=TRUE)
coalRelated <- grepl("coal", SCC$SCC.Level.Four, ignore.case=TRUE) 
coalCombustion <- (combustionRelated & coalRelated)
combustionSCC <- SCC[coalCombustion,]$SCC
combustionNEI <- NEI[NEI$SCC %in% combustionSCC,]
```

```{r ggp2}
ggp <- ggplot(combustionNEI,aes(factor(year),Emissions/10^5)) +
  geom_bar(stat="identity",fill="grey",width=0.75) +
  theme_bw() +  guides(fill=FALSE) +
  labs(x="year", y=expression("Total PM"[2.5]*" Emission (10^5 Tons)")) + 
  labs(title=expression("PM"[2.5]*" Coal Combustion Source Emissions Across US from 1999-2008"))

print(ggp)

```

Emissions from coal combustion related sources have decreased from 6 * 10^6 to below 4 * 10^6 from 1999-2008.

## Question 5
Subset the motor vehicles.
```{r vehicles}
vehicles <- grepl("vehicle", SCC$SCC.Level.Two, ignore.case=TRUE)
vehiclesSCC <- SCC[vehicles,]$SCC
vehiclesNEI <- NEI[NEI$SCC %in% vehiclesSCC,]
baltimoreVehicles <- vehiclesNEI[vehiclesNEI$fips==24510,]
```

Print the plot.
```{r ggp3}
ggp <- ggplot(baltimoreVehicles,aes(factor(year),Emissions)) +
  geom_bar(stat="identity",fill="grey",width=0.75) +
  theme_bw() +  guides(fill=FALSE) +
  labs(x="year", y=expression("Total PM"[2.5]*" Emission (10^5 Tons)")) + 
  labs(title=expression("PM"[2.5]*" Motor Vehicle Source Emissions in Baltimore from 1999-2008"))

print(ggp)
```

In Baltimore City, the emissions from motor vehicle has decreased from 1999-2008.

## Question 6
```{r vehiclesBaltimore}
vehiclesBaltimore <- vehiclesNEI[vehiclesNEI$fips == 24510,]
vehiclesBaltimore$city <- "Baltimore City"
vehiclesLANEI <- vehiclesNEI[vehiclesNEI$fips=="06037",]
vehiclesLANEI$city <- "Los Angeles County"
bothNEI <- rbind(vehiclesBaltimore,vehiclesLANEI)
```

```{r ggp4}
ggp <- ggplot(bothNEI, aes(x=factor(year), y=Emissions, fill=city)) +
  geom_bar(aes(fill=year),stat="identity") +
  facet_grid(scales="free", space="free", .~city) +
  guides(fill=FALSE) + theme_bw() +
  labs(x="year", y=expression("Total PM"[2.5]*" Emission (Kilo-Tons)")) + 
  labs(title=expression("PM"[2.5]*" Motor Vehicle Source Emissions in Baltimore & LA, 1999-2008"))

print(ggp)
```

It was shown in the plot that Los Angeles had a great change in motor vehicle emissions.